IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-07-07 paper/retrofit.tex
OK I think I socked the new intro
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

Our overall plan will be to take a \textbf{{\color{red} target
language}} with a formal operational semantics and combine it with an
\textit{{\color{blue} information flow control language}}.  For example,
taking ECMAScript as the target language and combining it with our IFC
language should produce the formal semantics for SWAPI.  In this
presentation, we use a simple, untyped lambda calculus with mutable
references and fixpoint in place of ECMAScript, to demonstrate some of the key
properties without having too much detail; we discuss the proper
embedding in more detail in Section~\ref{sec:real}.

@@

Most modern, dynamic information flow control languages encode
policy by associating a \emph{label}
with data, which encodes whether or not data tagged with one
label |l1| can flow to another label |l2| (written as |canFlowTo2 l1 l2|).
Our embedding is agnostic to the choice of labeling scheme; we only
require the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056}
with the partial order $\sqsubseteq$, join |lub|, and meet |glb|.
In this paper, we simply represent labels with the metavariable
$l$, but do not discuss them in more detail.

As opposed to more traditional fine-grained language-based IFC systems that
associate labels with values~\cite{sabelfeld2003language,
Austin:Flanagan:PLAS09}, we follow~\cite{Zeldovich:2006, lio} in taking a
coarse-grained floating-label approach and associating labels with
\emph{tasks}.
%
The task label---we refer to the label of the currently executing task as the
\emph{current label}---serves to protect everything in the task's scope;
all data in a task shares this common label.
%
Hence before performing a read or a write, the IFC monitor inspects the
current label to decide whether the operation is permitted.
%
A task can only write to entities that are at least as sensitive.
%
Similarly, it can only read from entities that are less sensitive.
%
However, as in other floating-label systems, this current label can be raised
to allow the task to read from more sensitive entities at the cost of giving up
the ability to write to certain entities.

>>>>

Information flow control systems operate by associating data with \emph{labels},
which specify whether or not data tagged with one label |l1| can flow to another
label |l2| (written as |canFlowTo2 l1 l2|).  These labels encode the desired
security policy (for example, confidential information should not flow to
a public channel), and the work of specifying the semantics of an information
flow language involves demonstrating that impermissible flows cannot happen,
a property called \emph{noninterference}.

Supporting fine-grained labeling of data in a programming language often
requires extremely intrusive modifications to its semantics and
implementation.  In this paper, we advocate applying coarse-grained
labeling~\cite{Zeldovich:2006, lio} to its extreme: our starting point
is multiple instances of completely isolated language runtimes or \emph{tasks}, with
information flow control applied between their communication.  For
languages which are implemented with isolation, such as JavaScript or C
with access to hardware privilege separation, systems along these lines
are far more easily implementable and require minimal semantic changes.

Why is this setting more interesting than ``traditional
coarse-grained information flow control system on isolated
threads?''  First, although completely separate tasks
are semantically easy to reason about, real-world implementations often blur the
lines between tasks in the name of efficiency.  Characterizing what optimizations
are permissible is a problem we address in Section~\ref{sec:concrete}, and
it is quite subtle, since removing transitions from the operational semantics
of a language can break noninterference.  Second, even though the underlying
semantic model involves communicating tasks, these tasks can be coordinated
together in ways that simulate traditional features in monolithic languages.
In fact, it is a very useful design tool for discovering what variants of the
features are permissible and which are not.

A very simple example is a low security task sending a high security
task a mutable reference:
\begin{align*}
    & |let i = TI (sandbox (brecvf x _ (IT (!(TI x)))))|\\
    & |in TI (send (IT i) getLabel (IT (ref true)))|
\end{align*}

For now, ignore |TI| and |IT|: roughly, this code creates a new
|sandbox|ed thread $i$ which waits (\textbf{blockingRecv}) for a
message, and then \textbf{send}s the thread a mutable reference to (|ref
true|).  If this operation actually shared the mutable cell, this cell
could be used to violate information flow control.  When working with
isolated tasks which each have their own heaps, the isolation is obviously true:
the sandboxed task receives a dangling pointer.

While this is a very simple example, we can still ask: given the
obviously safe formulation with separate heaps, can this fact be useful
for showing the safety of an semantics which only has a single heap
(Yes: see Section~\ref{sec:single-heap})?  Furthermore, the ``raw''
semantics violates expectations about how inter-thread communication
works on a shared heap: it strongly suggests, for example, that which
types of data may be passed between tasks should be restricted (for
example, in JavaScript, one might only allow JSON objects to be passed
between tasks).

Before moving on to the formal treatment of our system, we'd like to give
two more examples of language features which have an interesting interpretation
when viewed in the setting of multiple tasks.

\paragraph{Stack frames versus scheduling}  One feature which is
supported by some information flow control systems is the ability to
temporarily raise the current label in order to do computation on data
(as done by LIO's |toLabeled| and Breeze's
brackets~\cite{Hritcu:2013:YIB:2497621.2498098,
stefan:2012:arxiv-flexible}).  After the computation, the label is lowered back to the original label,
enforcing a stack discipline.  This is the trickiest part of the feature: care must be taken to ensure that information doesn't leak when
the label is lowered. In fact, systems with this feature generally only
achieve termination-insensitive non-interference: if the high secrecy
section doesn't terminate, this is observable by the non-execution of
the defunct public continuation.

\begin{figure}
\begin{align*}
    & |toLabeled l te @= | \\
    & |tab let i = taskId in| \\
%%    & |tab let j = TI (sandbox (IT (TI (setLabel l); TI (send i l (IT te)))))| \\
    & |tab let j = TI (sandbox (IT (TI (send i l (IT te)))))| \\
    & |tab in (\ _ . TI (setLabel l); TI (recvf x _ x unit))|
\end{align*}
\caption{Encoding of toLabeled using isolated tasks.}\label{fig:toLabeled}
\end{figure}

When considering tasks, we will instead say that the allocation of a new
stack frame with a raised label corresponds to the creation of a new
isolated task; a sample encoding using our language is shown in
Figure~\ref{fig:toLabeled}.
However, this is not enough: in a stack discipline, one generally
expects stack entries to be processed in order, rather than in parallel
as would be suggested by isolated tasks.  The sequential processing of
the task queue corresponds to a specific \emph{scheduling policy} of
these threads.  The recasting of the problem to one of scheduling
highlights the essential difference between termination-insensitive and
termination-sensitive non-interference.

\paragraph{Exceptions versus message passing}
Another language feature that has caused difficulty for
many information flow control systems is exceptions handling,
when built on top of this stack discipline.
%
The key difficulty appears when considering the propagation of exceptions
across stack boundaries where the current label is restored.
%
If the exception crosses this boundary, it can leak information about the
inner secret computation into a public computation; hence, one must
explicitly ensure that |toLabeled| does not propagate exceptions.

When stack frames are treated as isolated tasks, the correct semantics
are achieved automatically.
%
Exceptions simply stop when the hit the top-level of a task:
the only way for the frame to communicate information is by
initiating inter-task message passing.
%
For instance, 
\begin{code}
TI(toLabeled l (IT (if te then fail else true))); true
\end{code}
always evaluates to |true|, regardless of the value of |te|.
\Red{TODO: note that if the target language allows inspecting if
  result is exception we can implement Breeze's NAV (vs. just
propagating)}

\subsection{Preliminaries}

Our goal now is to
describe how to take a \textbf{{\color{red} target
language}} with a formal operational semantics and combine it with an
\textit{{\color{blue} information flow control language}}.  For example,
taking ECMAScript as the target language and combining it with our IFC
language should produce the formal semantics for SWAPI.  In this
presentation, we use a simple, untyped lambda calculus with mutable
references and fixpoint in place of ECMAScript, to demonstrate some of the key
properties without having too much detail (and because the embedding really
doesn't care about what, exactly, is in the target language); we discuss the proper
embedding in more detail in Section~\ref{sec:real}.

@@

As mentioned previously, most modern, dynamic information flow control
languages encode policy by associating a label with data.  Our
embedding is agnostic to the choice of labeling scheme; we only require
the labels to form a lattice~\cite{Denning:1976:LMS:360051.360056} with
the partial order $\sqsubseteq$, join |lub|, and meet |glb|.  In this
paper, we simply represent labels with the metavariable $l$, but do not
discuss them in more detail.

In our coarse-grained floating-label approach, labels are associated with tasks.
%
The task label---we refer to the label of the currently executing task as the
\emph{current label}---serves to protect everything in the task's scope;
all data in a task shares this common label.
%
Hence before performing a read or a write, the IFC monitor inspects the
current label to decide whether the operation is permitted.
%
A task can only write to entities that are at least as sensitive.
%
Similarly, it can only read from entities that are less sensitive.
%
However, as in other floating-label systems, this current label can be raised
to allow the task to read from more sensitive entities at the cost of giving up
the ability to write to certain entities.
